Structured matrix

### 2020-08-12 doc/fast_ssa/paper.tex
sync the content

The above operation can be expressed into the matrix-vector form. To simplify the definition of Fourier transform matrix, we define a variable $z = e^{-i \pi  / N}$, so the Fourier transform matrix, which is a $N$-by-$N$ square matrix and the element $F_{kl} = z^{kl}$. It can be expanded as
\begin{equation}
{\bf F}_N = \begin{bmatrix}
1 & 1 & \cdots & 1 \\
1 & z & \cdots & z^{N-1} \\
1 & z^2 & \cdots & z^{2(N-1)} \\
\vdots & \vdots & \ddots & \vdots \\
1 & z^{N-1} & \cdots & z^{(N-1)(N-1)} \\
\end{bmatrix}
\label{eq5}
\end{equation}
As ${\bf F}_N$ is a orthogonal (but not orthonormal) matrix, we can be easily get the its inverse matrix
\begin{equation}
{\bf F}_N^{-1} = \frac{1}{N} {\bf F}_N^H = \frac{1}{N} {\bf F}_N^{\ast} \,
\label{eq6}
\end{equation}
where the super-script $^H$ represent complex conjugate transpose and $\ast$ indicate conjugate transpose. As the matrix ${\bf F}_N$ is a symmetrical matrix, that's why we can get equation \ref{eq6}.

@@

\subsection{Circulant matrix}
An $9 \times 9$ circulant matrix is defined as
\begin{equation}
{\bf C} = 
\begin{bmatrix}
c_1 & c_9 & c_8 & c_7 & c_6 & c_5 &  c_4 & c_3 & c_2 \\ 
c_2 & c_1 & c_9 & c_8 & c_7 & c_6 &  c_5 & c_4 & c_3 \\ 
c_3 & c_2 & c_1 & c_9 & c_8 & c_7 &  c_6 & c_5 & c_4 \\ 
c_4 & c_3 & c_2 & c_1 & c_9 & c_8 &  c_7 & c_6 & c_5 \\ 
c_5 & c_4 & c_3 & c_2 & c_1 & c_9 &  c_8 & c_7 & c_6 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 & c_1 &  c_9 & c_8 & c_7 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 & c_2 &  c_1 & c_9 & c_8 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 & c_3 &  c_2 & c_1 & c_9 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 & c_4 &  c_3 & c_2 & c_1 \\ 
\end{bmatrix}
\end{equation} 
A circulant matrix is fully specified by its first column ${\bf c} = \begin{bmatrix}
c_1 & c_2 &  c_3 &  c_4 &  c_5 &  c_6 &  c_7 &  c_8 &  c_9  \end{bmatrix}^T$, we know that the product between circulant matrix and a vector ${\bf v}$ length of $N$ is equivalent to the circular convolution between vector ${\bf c}$ and ${\bf v}$, and the circular convolution can be computed efficiently via Fast Fourier transform (FFT). So we have
\begin{equation}
{\bf Cv} = {\bf F}_N^{-1} \left(  {\bf F}_N {\bf c} \circ {\bf F}_N {\bf v} \right)  =  {\bf F}_N^{-1} \left(  diag({\bf F}_N {\bf c}) \cdot {\bf F}_N {\bf v} \right) \,,
\end{equation}
where the symbol $\circ$ and $diag$ represents Hardmard (element-wise) multiplication and building diagonal matrix from a vector.

>>>>

The above operation can be expressed into the matrix-vector form. To simplify the definition of Fourier transform matrix, we define a variable $z = e^{-i 2\pi  / N}$, so the Fourier transform matrix, which is a $N$-by-$N$ square matrix and the element $F_{kl} = z^{kl}$. It can be expanded as
\begin{equation}
{\bf F}_N = \frac{1}{\sqrt{N}}\begin{bmatrix}
1 & 1 & \cdots & 1 \\
1 & z & \cdots & z^{N-1} \\
1 & z^2 & \cdots & z^{2(N-1)} \\
\vdots & \vdots & \ddots & \vdots \\
1 & z^{N-1} & \cdots & z^{(N-1)(N-1)} \\
\end{bmatrix}
\label{eq5}
\end{equation}
As ${\bf F}_N$ is a orthogonal (but not orthonormal) matrix, we can be easily get the its inverse matrix
\begin{equation}
{\bf F}_N^{-1} = {\bf F}_N^H =  {\bf F}_N^{\ast} \,
\label{eq6}
\end{equation}
where the super-script $^H$ represent complex conjugate transpose and $\ast$ indicate complex conjugate to each element. As the matrix ${\bf F}_N$ is a symmetrical matrix, that's why we can get equation \ref{eq6}.

@@

\subsection{   matrix}
An $9 \times 9$ circulant matrix is defined as
\begin{equation}
{\bf C} = 
\begin{bmatrix}
c_1 & c_9 & c_8 & c_7 & c_6 & c_5 &  c_4 & c_3 & c_2 \\ 
c_2 & c_1 & c_9 & c_8 & c_7 & c_6 &  c_5 & c_4 & c_3 \\ 
c_3 & c_2 & c_1 & c_9 & c_8 & c_7 &  c_6 & c_5 & c_4 \\ 
c_4 & c_3 & c_2 & c_1 & c_9 & c_8 &  c_7 & c_6 & c_5 \\ 
c_5 & c_4 & c_3 & c_2 & c_1 & c_9 &  c_8 & c_7 & c_6 \\ 
c_6 & c_5 & c_4 & c_3 & c_2 & c_1 &  c_9 & c_8 & c_7 \\ 
c_7 & c_6 & c_5 & c_4 & c_3 & c_2 &  c_1 & c_9 & c_8 \\ 
c_8 & c_7 & c_6 & c_5 & c_4 & c_3 &  c_2 & c_1 & c_9 \\ 
c_9 & c_8 & c_7 & c_6 & c_5 & c_4 &  c_3 & c_2 & c_1 \\ 
\end{bmatrix}
\end{equation} 
A circulant matrix is fully specified by its first column ${\bf c} = \begin{bmatrix}
c_1 & c_2 &  c_3 &  c_4 &  c_5 &  c_6 &  c_7 &  c_8 &  c_9  \end{bmatrix}^T$, we know that the product between circulant matrix and a vector ${\bf v}$ length of $N$ is equivalent to the circular convolution between vector ${\bf c}$ and ${\bf v}$, and the circular convolution can be computed efficiently via Fast Fourier transform (FFT). So we have
\begin{equation}
{\bf Cv} = {\bf F}_N^{-1} \left(  \sqrt{N}{\bf F}_N {\bf c} \circ {\bf F}_N {\bf v} \right)  =  {\bf F}_N^{-1} \left( \sqrt{N} diag({\bf F}_N {\bf c}) \cdot {\bf F}_N {\bf v} \right) \,,
\label{circ_times_vector}
\end{equation}
where the symbol $\circ$ and $diag$ represents Hardmard (element-wise) multiplication and building diagonal matrix from a vector.
Here we use a $4$-point vector to explain the above property, we define a vector ${\bf c} = \begin{bmatrix} c_1 & c_2 &c_3 & c_4\end{bmatrix}^T$. The circulant matrix formed by this vector is given as
\begin{equation}
{\bf C} = 
\begin{bmatrix}
c_1 & c_4 & c_3 & c_2  \\ 
c_2 & c_1 & c_4 & c_3  \\ 
c_3 & c_2 & c_1 & c_4  \\ 
c_4 & c_3 & c_2 & c_1 
\end{bmatrix}
\end{equation} 
Follow above definitions, we give $z = e^{-i 2\pi / 4}$, so the corresponding Fourier transform matrix is given as
\begin{equation}
{\bf F}_4 = \frac{1}{2}\begin{bmatrix}
1 & 1 & 1 & 1 \\
1 & z & z^2 & z^3 \\
1 & z^2 & z^4 & z^6 \\
1 & z^3 & z^6 & z^9 \\
\end{bmatrix}
\end{equation}
The inverse Fourier transform matrix is the complex conjugate transpose as
\begin{equation}
{\bf F}_4^H = \frac{1}{2}\begin{bmatrix}
1 & 1 & 1 & 1 \\
1 & z^{-1} & z^{-2}  & z^{-3}  \\
1 & z^{-2}  & z^{-4}  & z^{-6}  \\
1 & z^{-3}  & z^{-6}  & z^{-9}  \\
\end{bmatrix}
\end{equation}
The multiplication between the circulant matrix and the first column of ${\bf F}_4^H$ is given as
\begin{equation}
{\bf C}{\bf f}_1 = \frac{1}{2}
\begin{bmatrix}
c_1 & c_4 & c_3 & c_2  \\ 
c_2 & c_1 & c_4 & c_3  \\ 
c_3 & c_2 & c_1 & c_4  \\ 
c_4 & c_3 & c_2 & c_1 
\end{bmatrix}
\begin{bmatrix}
1 \\
1 \\
1 \\
1 
\end{bmatrix}
= \frac{c_1+c_2+c_3+c_4}{2}
\begin{bmatrix}
1 \\
1 \\
1 \\
1 
\end{bmatrix}
\end{equation} 
The coefficient in front of the vector is equation to the first component of the Fourier transform of vector ${\bf c}$ times $\sqrt{N}$, where $N$ is the total length of the vector.
Let's continue the discussion and multiply the circulant matrix by the second column of ${\bf F}_4^H$. We get
\begin{equation}
{\bf C}{\bf f}_1 = \frac{1}{2}
\begin{bmatrix}
c_1 & c_4 & c_3 & c_2  \\ 
c_2 & c_1 & c_4 & c_3  \\ 
c_3 & c_2 & c_1 & c_4  \\ 
c_4 & c_3 & c_2 & c_1 
\end{bmatrix}
\begin{bmatrix}
1 \\
z^{-1} \\
z^{-2} \\
z^{-3} 
\end{bmatrix}
= \frac{1}{2}
\begin{bmatrix}
c_1 + c_4 z^{-1} + c_3 z^{-2} + c_2 z^{-3}  \\
c_2 + c_1 z^{-1} + c_4 z^{-2} + c_3 z^{-3}  \\
c_3 + c_2 z^{-1} + c_1 z^{-2} + c_4 z^{-3}  \\
c_4 + c_3 z^{-1} + c_2 z^{-2} + c_1 z^{-3}  \\
\end{bmatrix}
= \frac{1}{2}
\begin{bmatrix}
1\left(c_1 + c_4 z^{-1} + c_3 z^{-2} + c_2 z^{-3} \right)  \\
z^{-1} \left(c_2 z    + c_1 + c_4 z^{-1} + c_3 z^{-2} \right)  \\
z^{-2} \left(c_3 z^2+ c_2 z + c_1 + c_4 z^{-1} \right)  \\
z^{-3} \left(c_4 z^3+ c_3 z^{-2} + c_2 z + c_1 \right) \\
\end{bmatrix}
\end{equation}
The function $z^n$ is a periodic function with period of $N=4$, the vector on the right hand side is equal to
\begin{equation}
\frac{1}{2}
\begin{bmatrix}
1\left(c_1 + c_4 z^{-1} + c_3 z^{-2} + c_2 z^{-3} \right)  \\
z^{-1} \left(c_2 z    + c_1 + c_4 z^{-1} + c_3 z^{-2} \right)  \\
z^{-2} \left(c_3 z^2+ c_2 z + c_1 + c_4 z^{-1} \right)  \\
z^{-3} \left(c_4 z^3+ c_3 z^{-2} + c_2 z + c_1 \right) \\
\end{bmatrix}
=
\frac{1}{2}
\begin{bmatrix}
1         \left(c_1 + c_2 z + c_3 z^2 + c_4 z^3 \right)  \\
z^{-1} \left(c_1 + c_2 z + c_3 z^2 + c_4 z^3 \right)  \\
z^{-2} \left(c_1 + c_2 z + c_3 z^2 + c_4 z^3 \right)  \\
z^{-3} \left(c_1 + c_2 z + c_3 z^2 + c_4 z^3 \right) \\
\end{bmatrix} 
= 
\frac{1}{2}\left(c_1 + c_2 z + c_3 z^2 + c_4 z^3 \right) 
\begin{bmatrix}
1         \\
z^{-1} \\
z^{-2} \\
z^{-3} \\
\end{bmatrix} 
\end{equation}
The define the scalar in front of the right vector as $\lambda = \frac{1}{2}\left(c_1 + c_2 z + c_3 z^2 + c_4 z^3 \right) $, which is the second element of the Fourier transform of the vector $c$ times $\sqrt{N}$. Following the same derivation, we can get
\begin{equation}
{\bf C} {\bf f}_i = \lambda_i  {\bf f}_i
\end{equation}
If we organize all the vectors into a matrix, we can get
\begin{equation}
{\bf C}{\bf F}^H = {\bf F}^H diag(\sqrt{N}{\bf F}{\bf c})
\end{equation}
We can further right multiply both end of the equations with ${\bf F}$, we can get
\begin{equation}
{\bf C} = {\bf F}^H diag(\sqrt{N}{\bf F}{\bf c}) {\bf F}
\label{decom}
\end{equation}
Which is equivalent to the expression in equation \ref{circ_times_vector}. In reality, we never formulate the Fourier transform matrix explicitly, instead the fast Fourier transform operator $fft$ and the inverse transform operator $ifft$ are used. we know
\begin{equation}
\begin{split}
\sqrt{N} {\bf F}_n = fft \\
 {\bf F}_N^H = \sqrt{N} ifft
 \end{split}
\end{equation}
with the above expression, we can replace the equation \ref{circ_times_vector} with the two operator as
\begin{equation}
{\bf C} {\bf v} = \sqrt{N} ifft\left( diag(fft({\bf c}))  {\frac{1}{\sqrt{N}} fft({\bf v})} \right)
\label{decom}
\end{equation}
With some simple manipulations, we can get the final form as
\begin{equation}
{\bf C} {\bf v} =  ifft \left( fft({\bf c}) \circ  fft({\bf v}) \right)
\label{decom}
\end{equation}
